---
title: "A Bayesian VAR Suite for Forecasting the Australian Economy"
subtitle: "Block-exogenous small-open-economy BVARs, pooled density forecasts, 1–12 quarters ahead"
date: today
format:
  html:
    toc: true
    embed-resources: true
    fig-align: center
execute:
  echo: false
  warning: false
---

```{r setup}
suppressPackageStartupMessages({library(knitr)})
tabdir <- "../output/tables"; figdir <- "../output/figures"
rd <- function(f) read.csv(file.path(tabdir, f), stringsAsFactors = FALSE)
spec    <- rd("transform_spec.csv")
scores  <- rd("scores_by_horizon.csv")
diag    <- rd("diagnostics.csv")
wfinal  <- rd("final_weights.csv")
dm      <- tryCatch(rd("dm_tests.csv"), error = function(e) NULL)
fc      <- read.csv("../output/forecasts/forecast_table.csv")
src <- yaml::read_yaml("../config/config.yml")$default$data$source
src_note <- if (identical(src, "synthetic")) {
  "All results below are produced from the synthetic block-exogenous DGP, which exists so that the pipeline runs end-to-end offline; the *relative* model rankings are informative about the machinery, not about Australia. Re-run with `data.source: real` for substantive results."
} else {
  "Results are based on final-vintage published data; real-time vintages are not used, so revisions-sensitive comparisons (especially for GDP) should be read with care."
}
```

## 1. What this is

This report documents a suite of Bayesian VARs for forecasting the key
Australian macroeconomic aggregates 1–12 quarters ahead, combined into a single
pooled **density** forecast. Australia is modelled as a **small open economy**:
every VAR member partitions the system into a foreign block (world activity,
commodity prices, the foreign policy rate) and a domestic block, and **block
exogeneity** forbids domestic variables from feeding back into the foreign
block (Cushman & Zha 1997; Zha 1999), at lags and contemporaneously. This is
imposed in estimation (not just on impulse responses) and verified numerically
(§6).

**Data source for this run:** `r src`. `r src_note`

## 2. Data and transformations

The `transform_spec` table is the single source of truth mapping each series to
its transform, Minnesota own-lag prior mean (`delta`), block, and steady-state
prior anchor (used by the Villani members).

```{r}
kable(spec[, c("variable","label","block","transform","delta","ss_mean","ss_sd")],
      digits = 2)
```

Growth-type series enter as 100·Δlog with `delta = 0`; rates and 100·log
levels enter with `delta = 1` plus sum-of-coefficients and
dummy-initial-observation priors in selected members to discipline the I(1)
behaviour at long horizons.

## 3. The suite

Seven block-exogenous VARs spanning prior family (Minnesota vs steady-state),
size (8-variable SOE core vs 13-variable medium), tightness (GLP-style
marginal-likelihood-selected λ vs fixed tight/loose), lag length (2/4/5) and
volatility (constant vs stochastic), plus four univariate anchors (random walk,
AR(4), UCSV, unconditional mean). All members are estimated recursively at
every forecast origin and produce predictive **draws** for every target
variable and horizon, so they pool cleanly.

## 4. Evaluation design

Pseudo-real-time expanding-window evaluation: at each origin *t* every member
sees only data through *t* — including the data-driven tightness selection,
which is re-run inside each origin. Combination weights at *t* use only scores
whose realization was observable at *t* (`s + h ≤ t`). Both properties are
enforced structurally and verified by automated tests (§6). Densities are
scored with the log score and CRPS (`scoringRules`), points with RMSE, and
calibration with PIT histograms, all by horizon, for quarterly and year-ended
transforms.

## 5. Results

### 5.1 Density accuracy by horizon

Mean CRPS by horizon (lower is better), members vs pooled combinations:

```{r, results="asis"}
for (v in unique(fc$variable)) {
  f <- file.path(figdir, sprintf("crps_by_h_%s.png", v))
  if (file.exists(f)) cat(sprintf("![](%s){width=85%%}\n\n", f))
}
```

```{r}
tab <- scores[scores$measure == "q" & scores$h %in% c(1, 4, 8, 12), ]
tab <- reshape(tab[, c("member","variable","h","crps")],
               idvar = c("member","variable"), timevar = "h", direction = "wide")
kable(tab[order(tab$variable, tab$crps.1), ], digits = 3,
      caption = "Mean CRPS at h = 1, 4, 8, 12 (quarterly transforms)")
```

```{r}
lsx <- scores[scores$measure == "q" & scores$h %in% c(1, 4, 8, 12), ]
lsx <- reshape(lsx[, c("member","variable","h","logdens")],
               idvar = c("member","variable"), timevar = "h", direction = "wide")
kable(lsx[order(lsx$variable, -lsx$logdens.1), ], digits = 2,
      caption = "Mean log predictive density at h = 1, 4, 8, 12 (higher = better)")
```

### 5.2 Significance

Diebold–Mariano tests (Harvey-corrected, Newey–West lag h−1) against the AR(4)
and random-walk benchmarks; stars on CRPS differentials (`***` 1%, `**` 5%,
`*` 10%).

```{r}
if (!is.null(dm) && nrow(dm)) {
  d1 <- dm[dm$h %in% c(1, 4, 8) & dm$measure == "q" &
           grepl("^combo_", dm$member), ]
  kable(d1[order(d1$variable, d1$h), c("member","reference","variable","h",
        "dm_crps","p_crps","sig_crps")], digits = 3, row.names = FALSE,
        caption = "DM tests: combinations vs univariate benchmarks")
}
```

### 5.3 Which models earn weight where

Final-origin combination weights (optimal pool, shrunk toward equal), by
variable and horizon bucket:

```{r}
wp <- aggregate(weight ~ member + bucket, data = wfinal[wfinal$scheme == "pool", ],
                FUN = mean)
wp <- reshape(wp, idvar = "member", timevar = "bucket", direction = "wide")
names(wp) <- sub("^weight\\.", "", names(wp))
kable(wp[order(-wp$near), ], digits = 3, row.names = FALSE,
      caption = "Optimal-pool weights averaged over target variables")
```

The BMA diagnostic (predictive-likelihood weights, no shrinkage) shows which
single member the data favour — by construction it concentrates and is **not**
the recommended combination:

```{r}
kable(aggregate(weight ~ member, data = wfinal[wfinal$scheme=="bma",], FUN = mean),
      digits = 3, caption = "BMA diagnostic weights (averaged over variables/buckets)")
```

```{r, results="asis"}
fs <- list.files(figdir, pattern = "^weights_pool_.*near", full.names = TRUE)
for (f in fs) cat(sprintf("![](%s){width=85%%}\n\n", f))
```

### 5.4 Calibration

PIT histograms by member (h = 1 and h = 4): a calibrated density forecast has
uniform PITs. Hump-shaped histograms indicate over-dispersion, U-shapes
under-dispersion.

```{r, results="asis"}
for (f in c("pit_h1.png", "pit_h4.png")) {
  p <- file.path(figdir, f)
  if (file.exists(p)) cat(sprintf("![](%s){width=95%%}\n\n", p))
}
```

### 5.5 Current forecasts

Fan charts from the optimal-pool combination at the final origin:

```{r, results="asis"}
fs <- list.files(figdir, pattern = "^fan_pool_", full.names = TRUE)
for (f in fs) cat(sprintf("![](%s){width=85%%}\n\n", f))
```

```{r}
cb <- fc[fc$member == "combo_pool", ]
kable(cb[cb$h %in% c(1, 2, 4, 8, 12), c("variable","h","point","lo90","hi90")],
      digits = 2, row.names = FALSE,
      caption = "Combined (optimal pool) forecasts: median and 90% interval")
```

## 6. Diagnostics

Section-9 self-checks, enforced as a hard gate in the pipeline:

```{r}
kable(diag, digits = 4)
```

- **Block exogeneity binds:** max |posterior-mean coefficient of a domestic lag
  inside a foreign equation| across all members and origins is
  `r format(max(diag$block_exog_max), digits = 2)` (engines with structural
  zeros report exactly 0).
- **No look-ahead:** corrupting all post-origin data leaves every forecast
  draw bit-identical.
- **Reproducibility:** re-running an origin with the same master seed gives
  identical draws.

## 7. Honest limitations

- **Final-vintage data.** The evaluation is pseudo-real-time on (for real runs)
  final data; genuine real-time vintages would change GDP-related rankings.
- **Combination gains are modest by design.** Equal weights are hard to beat
  (the combination puzzle); the payoff we target is *calibration and
  robustness*, not large point-forecast gains, and the tables should be read
  that way — where a single member beats the pools at a given horizon, the
  honest summary is that the pool buys insurance, not accuracy.
- **Conditional forecasts** use hard substitution, not Waggoner–Zha shock
  redistribution; they are a convenience hook, not a policy-scenario tool.
- **The synthetic DGP** is block-exogenous and Gaussian-with-outliers; it
  validates machinery (recovery of the block structure, calibration under a
  known truth), not Australian dynamics.
